NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE and PG lipids

### 2019-10-22 Manuscript/manuscriptPGPE.tex
UPdate text
Add density profile from GROMOS-CKP POPG simulatio with counterions only
Update text

@@

      POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0      & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0.15 \todoi{Concentration to be checked.}  & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 1.0  \todoi{Concentration to be checked.}  & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0      & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0.15 \todoi{Concentration to be checked.}  & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 1.0  \todoi{Concentration to be checked.}  & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\

@@

\subsection{Sodium binding to PE and PG lipid bilayers}
In our previous study about PS lipids \cite{antila18}, the underestimated increase of PC
headgroup order parameters upon addition of negaticely charged lipids
was related
to the overestimated counterion binding affinity, which would overcompensates the effect of negative charge.
Similar correlation is observed here in %other than GROMOS-CKP simulation:
Lipid 17 simulations, which exhibits stronger counter-ion binding affinity to pure POPG bilayer (Fig. \ref{CIdensPG})
and larger decrease in POPC $\beta$ carbon order parameter upon addition of POPG than other simulations.
Together with the lower area per molecule (59.5 {\AA}$^2$) than in experiments (66.1 {\AA}$^2$),
the results suggest that the counterions bind too strongly and shield the electrostatic repulsion between PG headgroups
in bilayers simulated with Lipid17 parameters.
In GROMOS-CKP simulations, however, the $\beta$-carbon order parameter of POPC
increases upon addition of POPG more than in CHARMM36 and Slipids simulations even though the binding affinities in these
simulations are similar. On the other hand, the responses of different hydrogens attached in $\alpha$-carbons
are different in all simulations except in Slipids, most sever in Lipid17 and GROMOS-CKP where other in increases
and other decreases \todo{Can we figure out the reason for this?}.
In conclusion, the counter-ion binding seems to be overestimated in Lipid17 with the strongest affinity among the
tested models, while with this data we cannot fully evaluate binding affinity in the other models.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/CIdensPG.eps}
  \caption{\label{CIdensPG}
    Counterion densities and area per lipids from simulations with PG lipids.
    Experimental area for POPG at 303 K is 66.1 {\AA}$^2$ and 67 {\AA}$^2$ for DPPC at 323 K \cite{pan12b}.
  }
  \todo{Data for GROMOS is also available without ions, but I could not analyze it, probably because tpr file was not correct.}
\end{figure}

>>>>

Zwitterionic PE is the second most abundant glycerophospholipid in eukaryotic cells
and has been related to the diseases \cite{vance15,calzada16,patel17}.
PE and PG affect membrane protein functionality \cite{hariharan18} and bind to various proteins \cite{yeagle14}.

@@

      POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0  & 250:250 & 18158 & 0   &  298  & 200 & 200 & \cite{CHARMM36POPCPOPG5050} \\ 
      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0.34 \todoi{Concentration calculated based in total amount of calcium ions. This may not be reasonable due to the lack of counterions.}  & 250:250 & 20798 & 128 &  298  & 200 & 200 & \cite{CHARMM36POPCPOPG5050150mMCaCl} \\
      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 1.36  \todoi{Concentration calculated based in total amount of calcium ions. This may not be reasonable due to the lack of counterions.} & 250:250 & 18114 & 445  &  298  & 200 & 200 & \cite{CHARMM36POPCPOPG50501000mMCaCl} \\
      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0  & 400:100 & 18664 & 0  &  298  & 200 & 200 & \cite{CHARMM36POPCPOPG4010} \\
      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 1.0 \todoi{Concentration calculated based in total amount of calcium ions. This may not be reasonable due to the lack of counterions.} & 400:100 & 18647 & 419  &  298  & 200 & 200 & \cite{CHARMM36POPCPOPG40101000mMCaCl} \\

@@

\subsection{Sodium binding to PE and PG lipid bilayers}
In our previous study about PS lipids \cite{antila18}, the underestimated increase of PC
headgroup order parameters upon addition of negaticely charged lipids
was related
to the overestimated counterion binding affinity, which would overcompensates the effect of negative charge.
Similar correlation is observed here in %other than GROMOS-CKP simulation:
Lipid 17 simulations, which exhibits stronger counter-ion binding affinity to pure POPG bilayer (Fig. \ref{CIdensPG})
and larger decrease in POPC $\beta$ carbon order parameter upon addition of POPG than other simulations.
Together with the lower area per molecule (59.5 {\AA}$^2$) than in experiments (66.1 {\AA}$^2$),
the results suggest that the counterions bind too strongly and shield the electrostatic repulsion between PG headgroups
in bilayers simulated with Lipid17 parameters.
In GROMOS-CKP simulations, however, the $\beta$-carbon order parameter of POPC
increases upon addition of POPG more than in CHARMM36 and Slipids simulations even though the binding affinities in these
simulations are similar. On the other hand, the responses of different hydrogens attached in $\alpha$-carbons
are different in all simulations except in Slipids, most sever in Lipid17 and GROMOS-CKP where other in increases
and other decreases \todo{Can we figure out the reason for this?}.
In conclusion, the counter-ion binding seems to be overestimated in Lipid17 with the strongest affinity among the
tested models, while with this data we cannot fully evaluate binding affinity in the other models.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/CIdensPG.eps}
  \caption{\label{CIdensPG}
    Counterion densities and area per lipids from simulations with PG lipids.
    Experimental area for POPG at 303 K is 66.1 {\AA}$^2$ and 67 {\AA}$^2$ for DPPC at 323 K \cite{pan12b}.
  }
\end{figure}
